---
title: "regression problem"
author: "dr. J"
date: "2/16/2020"
output: html_document
---

# Introduction:
Data are from the website http://fueleconomy.gov 
This is a problem in modeling/prediction.
The idea is this:  Fuel efficiency from 2010 and 2011 passenger cars and light trucks is provied.  Using data from 2010, find the best model to predict fuel effiency; fit the model using data from 2010, and evaluate model performance using 2011 data.  

Questions:  
1. Is there evidence of an improvemet in fuel efficiency from 2010 t 2011?  
2. What variables are most significant predictors of fuel efficienty?  
3. Suggest ways in which fuel efficienty could be improved in future years.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AppliedPredictiveModeling)
```

data set: fuel economy from US Gov't website

```{r}
data(FuelEconomy)
library(lattice)

```
### Plot shown in the text:
load the data
FE is fuel economy

```{r}
cars2010 <- cars2010[order(cars2010$EngDispl),]
cars2011 <- cars2011[order(cars2011$EngDispl),]
cars2010a <- cars2010
cars2010a$Year <- "2010 Model Year"
cars2011a <- cars2011
cars2011a$Year <- "2011 Model Year"
plotData <- rbind(cars2010a, cars2011a)
plotTheme <- bookTheme(FALSE)
plotTheme$plot.symbol$col <- rgb(.2, .2, .2, .5)
plotTheme$plot.symbol$cex <- 0.7
trellis.par.set(plotTheme)

xyplot(FE ~ EngDispl|Year, plotData,
xlab = "Engine Displacement",
ylab = "Fuel Efficiency (MPG)",
between = list(x = 1.2))

```

### variables in the data set:

```{r}
names(cars2010a)
```


### regession fit

```{r}
model1 = glm(FE~ EngDispl, data=cars2010a)
summary(model1)
plot(cars2010a$EngDispl, cars2010a$FE, xlab = 'Engine Displacement, L', ylab = 'Fuel Efficiency, MPG')
abline(model1, col='red')
```

### how does this fit model year 2011?
```{r}
plot(cars2011a$EngDispl, cars2011a$FE, xlab = 'Engine Displacement,L',ylab='MPG')
x = cars2011a$EngDispl
y = model1$coefficients[1] + model1$coefficients[2]*x
lines(x,y, col='red')
```



may benefit from a second order term

```{r}
model2 = glm(FE~ EngDispl + I(EngDispl^2), data = cars2010a)

summary(model2)
```

### include transmission in the equation:

```{r}
levels(cars2010a$Transmission)
model3 = glm(FE~ EngDispl+Transmission, data = cars2010a)
summary(model3)

```



### predictions
```{r}
plot(cars2011a$EngDispl, cars2011a$FE, ylab='MPG', xlab='Engine displacement, L')
newData = cars2011a
y = predict(model3, newdata = cars2011a)

lines(cars2011a$EngDispl,y, col='grey')

#fit a smooth line to y then plot
ySmooth= glm(y~cars2011a$EngDispl)
summary(ySmooth)
abline(ySmooth, col='red')

```

Transmission effects ??

```{r}

model3AOV = aov(cars2010a$FE~cars2010a$Transmission)
summary(model3AOV)
TukeyHSD(model3AOV)
```



