---
title: "regression problem"
author: "Jacky/Dottie"
date: "2/21/2020"
output: html_document
---

# Introduction:
Data are from the website http://fueleconomy.gov 
This is a problem in modeling/prediction.
The idea is this:  Fuel efficiency from 2010 and 2011 passenger cars and light trucks is provied.  Using data from 2010, find the best model to predict fuel effiency; fit the model using data from 2010, and evaluate model performance using 2011 data.  

Questions:  
1. Is there evidence of an improvement in fuel efficiency from 2010 to 2011?  
2. What variables are most significant predictors of fuel efficienty?  
3. Suggest ways in which fuel efficienty could be improved in future years.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AppliedPredictiveModeling)

library(dplyr)
```

data set: fuel economy from US Gov't website

```{r}
data(FuelEconomy)
library(lattice)

```
### Plot shown in the text:
load the data
FE is fuel economy

```{r}
cars2010 <- cars2010[order(cars2010$EngDispl),]
cars2011 <- cars2011[order(cars2011$EngDispl),]
cars2010a <- cars2010
cars2010a$Year <- "2010 Model Year"
cars2011a <- cars2011
cars2011a$Year <- "2011 Model Year"
plotData <- rbind(cars2010a, cars2011a)
plotTheme <- bookTheme(FALSE)
plotTheme$plot.symbol$col <- rgb(.2, .2, .2, .5)
plotTheme$plot.symbol$cex <- 0.7
trellis.par.set(plotTheme)

xyplot(FE ~ EngDispl|Year, plotData,
xlab = "Engine Displacement",
ylab = "Fuel Efficiency (MPG)",
between = list(x = 1.2))

```

### variables in the data set:

```{r}
names(cars2010a)
```

Try transmissions = AV, AM6, and AM7

```{r}
#levels(cars2010a$Transmission)

testTranny = filter(cars2010a, Transmission %in% c('AV','AM6'))

trannyModel = glm(FE ~ Transmission, data = testTranny)
summary(trannyModel)

mean(testTranny$FE)
print(mean(testTranny$FE))
```

AM7 is no different from AM6 because is not significant because of Pr(>|t|) = 0.211 which is greater than 0.05, so we remove it.

AM6 = AIC: 466.37, AM7 = AIC: 427.5, which means if you had to choose between 6-speed or 7-speed, it's better to go with a 7-speed transmission vehicle.


Try adding AirAspirationMethod to see what it does.

```{r}
#levels(cars2010a$AirAspirationMethod)

trannyAirModel = glm(FE ~ Transmission + AirAspirationMethod, data = testTranny)
summary(trannyAirModel)
```

AirAspirationMethodTurbocharged makes no significant difference.

Next, try ExhaustValvesPerCyl to see what happens.

```{r}
#levels(cars2010a$ExhaustValvesPerCyl)

trannyValveModel = glm(FE ~ Transmission + ExhaustValvesPerCyl, data = testTranny)
summary(trannyValveModel)
```

ExhaustValvesPerCyl does not make any difference because there are no astericks next to 0.07386

```{r}
for (i in 1:length(cars2010a$IntakeValvePerCyl)) {
  if (cars2010a$IntakeValvePerCyl[i] == 0) print('found one')
}

#count(cars2010a$IntakeValvePerCyl == 0)
```

There are 2.

## Question 1. Is there evidence of an improvement in fuel efficiency from 2010 to 2011? 

Pull out the fuel efficiency from 2010 and 2011 data set and compare the mean difference.

Need to perform a t-test. Null and alternative hypothesis.

$H_0$: no difference between FE in 2010 and 2011
$H_A$: there is a difference in FE between 2010 and 2011

We'll use a t-test for this inference:

```{r}
t.test(cars2010a$FE, cars2011a$FE)
```

# Answer to Quesetion 1:

There is no evidence/difference in Fuel Efficiency from 2010 data set to 2011 data set because of the high P-value.



### regession fit

```{r}
model1 = glm(FE~ EngDispl, data=cars2010a)
summary(model1)
plot(cars2010a$EngDispl, cars2010a$FE, xlab = 'Engine Displacement, L', ylab = 'Fuel Efficiency, MPG')
abline(model1, col='red')
```

### how does this fit model year 2011?
```{r}
plot(cars2011a$EngDispl, cars2011a$FE, xlab = 'Engine Displacement,L',ylab='MPG')
x = cars2011a$EngDispl
y = model1$coefficients[1] + model1$coefficients[2]*x
lines(x,y, col='red')
```



may benefit from a second order term

```{r}
model2 = glm(FE~ EngDispl + I(EngDispl^2), data = cars2010a)

summary(model2)
```

### include transmission in the equation:

```{r}
levels(cars2010a$Transmission)
model3 = glm(FE~ EngDispl+Transmission, data = cars2010a)
summary(model3)

```



### predictions
```{r}
plot(cars2011a$EngDispl, cars2011a$FE, ylab='MPG', xlab='Engine displacement, L')
newData = cars2011a
y = predict(model3, newdata = cars2011a)

lines(cars2011a$EngDispl,y, col='grey')

#fit a smooth line to y then plot
ySmooth= glm(y~cars2011a$EngDispl)
summary(ySmooth)
abline(ySmooth, col='red')

```

Transmission effects ??

```{r}

model3AOV = aov(cars2010a$FE~cars2010a$Transmission)
summary(model3AOV)
TukeyHSD(model3AOV)
```



